Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-06-09 paper/main.tex
latest versio for the journal

\begin{abstract}
We study the alignment of dark matter halos with the cosmic web
characterized by the tidal and velocity shear fields. We focus on the
alignment of their shape, angular momentum and peculiar velocities. We  
use a cosmological N-body simulation that allows to study dark matter
halos spanning almost five orders of magnitude in mass
($10^{9}$-$10^{14}$) \hMsun and spatial scales of
$(0.5$-$1.0)$\hMpc to define the cosmic web. We find that the halo shape
presents the strongest alignment along the smallest tidal eigenvector,
e.g. along filaments and walls, with a signal that gets stronger as
the halo mass increases. In the case of the velocity shear field only 
massive halos $>10^{12}$\hMsun tend to have their shapes aligned along
the largest tidal eigenvector; that is, perpendicular to filaments and
walls. For the angular momentum we find alignment signals only for
halos more massive than $10^{12}$\hMsun both in the tidal and velocity
shear webs where the preferences are for it to be parallel to the
middle eigenvector; perpendicular to filaments and parallel to
walls. Finally, the peculiar velocities show a strong alignment along
the smallest tidal eigenvector for all halo masses; halos move along filaments
and walls. In the velocity shear the same alignment is present but
weaker and only for halos less massive than $10^{12}\hMsun$. Our
results clearly show that the two different algorithms we used to
define the cosmic web describe different physical aspects of
non-linear collapse and should be used in a complementary way to
understand the effect of the cosmic web on galaxy evolution.    
\end{abstract}
\begin{keywords}
methods: numerical, galaxies: haloes, cosmology: theory, dark
matter, large-scale structure of Universe
\end{keywords}

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1$ &
$\hat{e}_3$ (filament) &$++$ & $>10^{12}$\hMsun\\
&   & &
$\hat{e}_3$ (filament) & $+$ & $<10^{12}$\hMsun\\

@@

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1$ &
$\hat{e}_3$ (filament) &$--$ & $>10^{12}$\hMsun\\
&   & &
$\hat{e}_3$ (filament) & none & $<10^{12}$\hMsun\\
&   & &
$\hat{e}_1$ (wall) & $--$ & $>10^{12}$\hMsun\\

@@

&   & &
$\hat{e}_3$  (filament) & none & $<10^{12}$\hMsun\\

>>>>

\begin{abstract}
We study the alignment of dark matter halos with the cosmic web
characterized by the tidal and velocity shear fields. We focus on the
alignment of their shape, angular momentum and peculiar velocities. We  
use a cosmological N-body simulation that allows to study dark matter
halos spanning almost five orders of magnitude in mass
($10^{9}$-$10^{14}$) \hMsun and spatial scales of
$(0.5$-$1.0)$\hMpc to define the cosmic web. The strongest alignment
is measured for halo shape along the smallest tidal eigenvector,
e.g. along filaments and walls, with a signal that gets stronger as
the halo mass increases. In the case of the velocity shear field only
massive halos $>10^{12}$\hMsun tend to have their shapes aligned along
the largest tidal eigenvector; that is, perpendicular to filaments and
walls. For the angular momentum we find alignment signals only for
halos more massive than $10^{12}$\hMsun both in the tidal and velocity
shear fields where the preferences are for it to be parallel to the
middle eigenvector; perpendicular to filaments and parallel to
walls. Finally, the peculiar velocities show a strong alignment along
the smallest tidal eigenvector for all halo masses; halos move along filaments
and walls. The same alignment is present with the velocity shear,
albeit weaker and only for halos less massive than $10^{12}\hMsun$. Our
results show that the two different algorithms we used to
define the cosmic web describe different physical aspects of
non-linear collapse and should be used in a complementary way to
understand the cosmic web influence on galaxy evolution.    
\end{abstract}
\begin{keywords}
methods: numerical, galaxies: haloes, cosmology: theory, dark
matter, large-scale structure of Universe
\end{keywords}

@@

{\bf Forero-Romero et al. (2014)} & T-Web & $0.5-1$ &
$\hat{e}_3$ (filament) &$++$ & $>10^{12}$\hMsun\\
{\bf (This Work)}&   & &
$\hat{e}_3$ (filament) & $+$ & $<10^{12}$\hMsun\\

@@

{\bf Forero-Romero et al. (2014)} & Vp-Web & $0.5-1$ &
$\hat{e}_3$ (filament) &$--$ & $>10^{12}$\hMsun\\
{\bf (This Work)}&   & &
$\hat{e}_3$ (filament) & none & $<10^{12}$\hMsun\\
&   & &
$\hat{e}_1$ (wall) & $--$ & $>10^{12}$\hMsun\\

@@

{\bf (This Work)}&   & &
$\hat{e}_3$  (filament) & none & $<10^{12}$\hMsun\\
